---
title: "09"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(tictoc)
library(sp)
library(sf)
options(scipen = 999)
```

## Part 1

```{r}
input <- "7,1
11,1
11,7
9,7
9,5
2,5
2,3
7,3"


input <- "input/09"

parse_coord <- function(r) {
  str_split_1(r, ',') |> as.numeric()
}

compute_area <- function(a, b) {
  (abs(a[1] - b[1]) + 1) * (abs(a[2] - b[2]) + 1)
}

tic()
combn(read_lines(input), 2, simplify = FALSE) |> 
  map(\(x) list(parse_coord(x[1]), parse_coord(x[2]))) |> 
  enframe() |> 
  mutate(c1 = map(value, 1),
         c2 = map(value, 2)) -> rectangles
mutate(rectangles, area = map2_dbl(c1, c2, compute_area)) |> 
  slice_max(area, n = 1)
toc()

```

4781377701

4.211 sec elapsed

## Part 2

max_x 98240
max_y 98457

Sort the area from part 1 from large to small and test if overlap with green areas.
As soon as one is valid, break.


```{r}
tibble(coord = read_lines(input)) |> 
  mutate(cc = map(coord, parse_coord)) |> 
  unnest_wider(col = cc, names_sep = "_") |> 
  rename(x = cc_1, y = cc_2) -> red_tiles

red_tiles
```
```
# A tibble: 28 × 7     x1    y1    x2    y2
    name value       c1_1  c1_2  c2_1  c2_2  area
   <int> <list>     <dbl> <dbl> <dbl> <dbl> <dbl>
 1    17 <list [2]>    11     7     2     3    50  (11 - 2 + 1) * ( 7 - 3 + 1)
      21 <list [2]>	    9	    7	    2	    3    40
```


Using fancy `{sp}` instead and smart **Polygon**

```{r}
select(red_tiles, x, y) |> 
  as.matrix() |> 
  Polygon() -> green_area

green_area <- SpatialPolygons(list(Polygons(list(green_area), ID = 1)))

tic()
mutate(rectangles, area = map2_dbl(c1, c2, compute_area)) |> 
  arrange(desc(area)) |> 
  unnest_wider(col = c(c1, c2), names_sep = "_") -> part2


areas <- vector(mode = "list", length = nrow(part2))

for (i in seq_len(nrow(part2))) {
  matrect <- tibble(x = c(part2$c1_1[i], part2$c2_1[i],
                          part2$c1_1[i], part2$c2_1[i]),
                    y = c(part2$c1_2[i], part2$c2_2[i],
                          part2$c2_2[i], part2$c1_2[i])) |> 
    arrange(x) |> 
    as.matrix()
  rect_sp <- SpatialPolygons(list(Polygons(list(Polygon(matrect)), ID = 1)))
  
  inters <- st_intersection(st_as_sf(green_area), st_as_sf(rect_sp))
  
  if (i %% 10000 == 0) message(paste(i, " rectangles evaluated so far, area: ", part2[i, "area"]))
  
  areas[[i]] <- c(st_area(st_as_sf(rect_sp)), st_area(st_as_sf(inters)))
  # rect fully contained
  if (st_area(st_as_sf(inters)) == st_area(st_as_sf(rect_sp))) {
    message(paste(i, " rectangles evaluated so far, area: ", part2[i, "area"]))
    break
  }
}
toc()



plot(green_area)
plot(rect_sp) # plot(rect_sp, add = TRUE) in console to superpose
```
4767548688 too high
4602347920 too high
1470584154 not correct
1470511006 not correct
1470616992 <- 

50831  rectangles evaluated so far, area:  1470616992 (sp computes (94821 - 5254)*(66490 - 50072) = 1470511006)



```{r}
part2 |> 
  mutate(ll = areas) |> 
  unnest_wider(ll, names_sep = "_") |> 
  filter(ll_1 == ll_2) |> 
  slice_max(ll_1, n = 1)
```




Playing with https://stackoverflow.com/a/62450167/1395352

```{r}
rectover <- structure(c(20000, 20000, 60000, 60000, 10000, 14000, 14000, 
                        10000), dim = c(4L, 2L), dimnames = list(NULL, c("x", "y")))
rectin <- structure(c(30000, 30000, 60000, 60000, 10000, 14000, 14000, 
                      10000), dim = c(4L, 2L), dimnames = list(NULL, c("x", "y")))
rect_sp <- SpatialPolygons(list(Polygons(list(Polygon(rectin)), ID = 1)))
interin = ll

inters <- st_intersection(st_as_sf(green_area), st_as_sf(rect_sp))

plot(green_area, col = "green3")
plot(rect_sp, col = "tomato2", add = TRUE)
plot(inters, add = TRUE, col = "orange")
```

List of 1
 $ : int 1
 - attr(*, "predicate")= chr "overlaps"
 - attr(*, "region.id")= chr "1"
 - attr(*, "remove_self")= logi FALSE
 - attr(*, "retain_unique")= logi FALSE
 - attr(*, "ncol")= int 1
 - attr(*, "class")= chr [1:2] "sgbp" "list"
