---
format: html
---


```{r}
#| label: setup
#| include: false

library(tidyverse)
```


## Part 1


- It contains at least **three vowels** (`aeiou` only), like aei, xazegov, or aeiouaeiouaeiou.
- It contains at least **one letter that appears twice in a row**, like `xx`, abcdde (dd), or aabbccdd (aa, bb, cc, or dd).
- It does **not** contain the strings `ab, cd, pq, or xy`, even if they are part of one of the other requirements.

```{r}
input <- read_lines("input/05")
s <- c(
  "ugknbfddgicrmopn",
  "aaa",
  "jchzalrnumimnmhp",
  "haegwjzuvuyypxyu",
  "dvszwmarrgswjxmb"
)

tibble(s = input) |>
  mutate(
    vowels = str_count(s, "[aeiou]") >= 3,
    repeated = str_detect(s, "([a-z])\\1"),
    forbidden = str_detect(s, "ab|cd|pq|xy", negate = TRUE),
    nice = (vowels + repeated + forbidden) == 3L
  ) |>
  summarise(sum(nice))
```


255


## Part 2

- It contains a pair of any two letters that appears at least twice in the string without overlapping, like `xyxy` (`xy`) or `aabcdefgaa` (`aa`), but not like `aaa` (`aa`, but it overlaps).
- It contains at least one letter which repeats with exactly one letter between them, like `xyx`, `abcdefeghi` (`efe`), or even `aaa`.


```{r}
input <- read_lines("input/05")
s <- c(
  "qjhvhtzxzqqjkmpb",
  "xxyxx",
  "uurcxstgmygtbstg",
  "ieodomkazucvgmuy",
  "xyxy",
  "aabcdefgaa",
  "aaa"
)

tibble(s = input) |>
  mutate(
    pairs = str_detect(s, "([a-z][a-z])[a-z]*\\1"),
    repeated = str_detect(s, "([a-z])[a-z]\\1"),
    nice = (pairs + repeated) == 2L
  ) |>
  summarise(sum(nice))
```

42 wrong
55